---
title: "DATA 608 - Module 1"
author: "Douglas Barley"
date: "9/5/2021"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```

```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r packages, include=FALSE}

# Insert your code here, create more chunks as necessary
library(tidyverse)
library(ggplot2)
```

``` {r cleanup_explore}
# remove NA values
inc <- na.omit(inc)

# explore the data
inc %>% 
  dplyr::select(-Name, -Industry, -City, -State, -Rank) %>% #remove categorical and Rank variable
    gather(key,value) %>%
      ggplot(aes(value)) + geom_boxplot() + facet_wrap(~key, scales = "free")
```

```{r explore_results}
# every numerical variable has many outliers, can we see why?

psych::describe(inc)

# there are considerable skews in the three numerical variables,
# as well as high kurtosis (i.e. number of outliers) in each variable

```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r Q1}

# Answer Question 1 here
inc_State <- data.frame(inc) %>%
  group_by(State) %>%
    summarize(
      state_count = n()
    ) %>%
      arrange(desc(state_count))

ggplot(data=inc_State, aes(x=reorder(State, -state_count), y=state_count)) +
  geom_col(fill="darkgrey") + coord_flip() +
  labs(x="State", y="Count", title="Companies by State")
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r 3rd_State}

# Answer Question 2 here
head(inc_State) # The state with the 3rd most companies is NY with 311

```

``` {r NY_data}
# gather the data for NY only from the original data

inc_NY_detail <-  data.frame(inc) %>%
  filter(State=="NY", complete.cases(.))

inc_NY <-  data.frame(inc) %>%
  filter(State=="NY", complete.cases(.)) %>%
    group_by(Industry) %>%
      summarize(
        tot_companies = n(),
        tot_emp = sum(Employees),
        avg_emp = mean(Employees),
        median_emp = median(Employees),
        min_emp = min(Employees),
        max_emp = max(Employees)
      ) %>%
        arrange(desc(tot_emp))  # sort in descending order by total workforce count in the 
                                # industry (i.e. industries with largest total workforce on top)

inc_NY
```

``` {R NY_graphic}

NY_plot <- ggplot(data=inc_NY_detail, aes(x=reorder(Industry,Employees,median), y=Employees)) +
            geom_boxplot() + coord_flip() +
            labs(x = "Industry", y = "Median # Employees", title="Median # Employees per Industry in NY State")

NY_plot + scale_y_log10()  # scale the y axis (x flipped axis) for better presentation

```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r investor_data}

# Answer Question 3 here
inc_invest <-  data.frame(inc) %>%
  filter(complete.cases(.)) %>%
    group_by(Industry) %>%
      summarize(
        tot_revenue = sum(Revenue),
        tot_emp = sum(Employees)
      ) %>%
        mutate(rev_per_emp = tot_revenue/tot_emp) %>%
          arrange(desc(rev_per_emp))

inc_invest
```

```{r investor_graphic}

Investor_plot <- ggplot(data=inc_invest, aes(x=reorder(Industry,rev_per_emp), y=rev_per_emp)) +
                  geom_bar(stat="identity", fill="darkgreen") + coord_flip() +
                  labs(x = "Industry", y = "Revenue per Employee",
                        title="Revenue per Employee by Industry") +
                  scale_y_continuous(limits=c(0,1450000)) +
                  geom_text(aes(label = round(rev_per_emp,0)), size=3, hjust=-0.25)

Investor_plot
```

